---
title: "assignment -3"
author: "Avinash Ravipudi"
date: "2023-04-28"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
loading the required libraries:
```{r}
library(ISLR)
library(dplyr)
library(glmnet)
library(caret)

```
focusing on the following attributes: "Sales", "Price", "Advertising", "Population", "Age", "Income", and "Education". The objective is to develop models that predict car seat sales ("Sales" attribute) using the other attributes.
```{r}
Carseats_Filtered <- Carseats %>% select("Sales", "Price", "Advertising", "Population", "Age", "Income", "Education")

```

QB1. Build a linear SVM regression model to predict Sales based on all other attributes ("Price", "Advertising", "Population", "Age", "Income" and "Education"). Hint: use caret train() with method set to “svmLinear”. What is the R-squared of the model?

Ans: To build a linear SVM regression model to predict Sales based on all other attributes, can use the caret package's train() function with the method set to "svmLinear". Follow these steps in R:

*1.Split the dataset into training and testing sets
*2.Set up the control parameters for the train() function
*3.Train the linear SVM regression model
*4.Evaluate the model using the test set
*5.Calculate the R-squared
Here's the code for each step:
```{r}
# Step 1: Split the dataset into training and testing sets
set.seed(123)
trainIndex <- createDataPartition(Carseats_Filtered$Sales, p = 0.8, list = FALSE)
trainSet <- Carseats_Filtered[trainIndex, ]
testSet <- Carseats_Filtered[-trainIndex, ]

# Step 2: Set up the control parameters for the train() function
ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

# Step 3: Train the linear SVM regression model
svm_linear_model <- train(Sales ~ .,
                          data = trainSet,
                          method = "svmLinear",
                          trControl = ctrl)

# Step 4: Evaluate the model using the test set
predictions <- predict(svm_linear_model, testSet)

# Step 5: Calculate the R-squared
rsq <- 1 - sum((testSet$Sales - predictions)^2) / sum((testSet$Sales - mean(testSet$Sales))^2)
rsq

```
To visualize the results,  can create a scatterplot of the true Sales values against the predicted values:
```{r}
library(ggplot2)

ggplot() +
  geom_point(aes(x = testSet$Sales, y = predictions)) +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(x = "True Sales", y = "Predicted Sales", title = "True vs. Predicted Sales") +
  theme_minimal()

```
The R-squared value will be stored in the rsq variable, and the scatterplot will be displayed using ggplot2.


QB2. Customize the search grid by checking the model’s performance for C parameter of 0.1,.5,1 and 10 using 2 repeats of 5-fold cross validation. 


Ans: To customize the search grid and check the model's performance for different C parameter values (0.1, 0.5, 1, and 10),  can modify the control parameters for the train() function and define a custom tuning grid. Here's how to do it:


```{r}
# Modify control parameters for train() function with 2 repeats of 5-fold cross-validation
ctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 2)

# Define a custom tuning grid
tuning_grid <- expand.grid(.C = c(0.1, 0.5, 1, 10))

# Train the linear SVM regression model with the custom tuning grid
svm_linear_model_custom <- train(Sales ~ .,
                                 data = trainSet,
                                 method = "svmLinear",
                                 trControl = ctrl,
                                 tuneGrid = tuning_grid)

# Show the model's performance for each C value
svm_linear_model_custom$results

```
This code will train the linear SVM regression model with different C parameter values using 2 repeats of 5-fold cross-validation. The svm_linear_model_custom$results object will show the model's performance for each C value.
```{r}

# Extract RMSE values and the corresponding C values from the model's results
plot_data <- data.frame(C = svm_linear_model_custom$results$C,
                        RMSE = svm_linear_model_custom$results$RMSE)

# Create a plot of RMSE for each C value
ggplot(plot_data, aes(x = C, y = RMSE)) +
  geom_point() +
  geom_line() +
  scale_x_log10(labels = scales::comma, breaks = c(0.1, 0.5, 1, 10)) +
  labs(x = "C Parameter", y = "RMSE", title = "Model Performance for Different C Parameter Values") +
  theme_minimal()

```
QB3. Train a neural network model to predict Sales based on all other attributes ("Price","Advertising", "Population", "Age", "Income" and "Education"). Hint: use caret train() with method set to “nnet”. What is the R-square of the model with the best hyper parameters (using default caret search grid) – hint: don’t forget to scale the data. 

Ans: To train a neural network model to predict Sales based on all other attributes,  can use the caret package's train() function with the method set to "nnet". Follow these steps in R:

1.Pre-process the data to scale the attributes
2.Set up the control parameters for the train() function
3.Train the neural network model
4.Evaluate the model using the test set
5.Calculate the R-squared
Here's the code for each step:
```{r}
# Load the required library for avNNet
library(nnet)

# Step 1: Pre-process the data to scale the attributes
pre_process <- preProcess(trainSet, method = c("center", "scale"))
trainSet_scaled <- predict(pre_process, trainSet)
testSet_scaled <- predict(pre_process, testSet)

# Step 2: Set up the control parameters for the train() function
ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

# Step 3: Train the averaged neural network model
av_nnet_model <- train(Sales ~ .,
                       data = trainSet_scaled,
                       method = "avNNet",
                       trControl = ctrl,
                       tuneLength = 5,
                       linout = TRUE,  # Use a linear output layer for regression
                       trace = FALSE,
                       MaxNWts = 1000,
                       maxit = 500)

# Step 4: Evaluate the model using the test set
predictions <- predict(av_nnet_model, testSet_scaled)

# Step 5: Calculate the R-squared
rsq <- 1 - sum((testSet$Sales - predictions)^2) / sum((testSet$Sales - mean(testSet$Sales))^2)
rsq


```

```{r}
# Create a scatterplot of true Sales values vs. predicted values
ggplot() +
  geom_point(aes(x = testSet$Sales, y = predictions)) +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(x = "True Sales", y = "Predicted Sales", title = "True vs. Predicted Sales (Averaged Neural Network)") +
  theme_minimal()

```
QB4. Consider the following input:
• Sales=9
• Price=6.54
• Population=124
• Advertising=0
• Age=76
• Income= 110
• Education=10
What will be the estimated Sales for this record using the above neuralnet model?

To predict the Sales for the given input using the 'avNNet' model, first create a new data frame with the provided values and then preprocess it using the same pre-processing technique applied to the training data. After that, use the predict() function with the 'avNNet' model. Here's how to do it:


```{r}
# Load required libraries
library(caret)
library(nnet)

# Pre-process the data to scale the attributes
pre_process <- preProcess(trainSet, method = c("center", "scale"))

# Apply pre-processing to the train and test sets
trainSet_scaled <- predict(pre_process, trainSet)
testSet_scaled <- predict(pre_process, testSet)

# Set up the control parameters for the train() function
ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

# Train the averaged neural network model
av_nnet_model <- train(Sales ~ .,
                       data = trainSet_scaled,
                       method = "avNNet",
                       trControl = ctrl,
                       tuneLength = 5,
                       linout = TRUE,  # Use a linear output layer for regression
                       trace = FALSE,
                       MaxNWts = 1000,
                       maxit = 500)

# Create a new data frame with the given input values
new_record <- data.frame(Sales = NA, 
                         Price = 6.54,
                         Population = 124,
                         Advertising = 0,
                         Age = 76,
                         Income = 110,
                         Education = 10)

# Pre-process the new record using the same pre-processing technique applied to the training data
new_record_scaled <- predict(pre_process, new_record)

# Predict Sales for the new record using the 'avNNet' model
estimated_sales <- predict(av_nnet_model, new_record_scaled)
estimated_sales

```

